---
title: "Estimating commodities and costs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating commodities and costs}
  %\VignetteEngine{rmarkdown::html_vignette}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette provides a brief example of estimating the number of
commodities required for a simple scenario and how these can be
converted into costs.  We use two small data sets as examples.

## Bed nets

First we consider some basic intervention information that includes
population at risk (`par`) and a target level of insecticide treated net
(ITN) usage by year.

```{r itn-data}
interventions <- data.frame(
  year    = 1:6,
  itn_use = c(0, 0.6, 0.5, 0.2, 0.5, 0.4),
  par     = rep(1000, 6)
)
interventions
```

To determine how many nets are required we can use `commodity_nets()`.
Here we assume a constant usage rate, distributions on the first day of
each year, and a half life of two years.

```{r nets}
dist_steps <- cumsum(c(1, rep(365, 5)))
crop_steps <- dist_steps + 183
half_life  <- 730
use_rate   <- 0.5

interventions$n_nets <- commodity_nets(
  usage                 = interventions$itn_use,
  use_rate              = use_rate,
  distribution_timesteps = dist_steps,
  crop_timesteps        = crop_steps,
  half_life             = half_life,
  par                   = interventions$par
)

interventions$cost_nets <- cost_llin(interventions$n_nets)

interventions
```

```{r plot-nets, fig.cap="Nets required by year"}
barplot(
  interventions$n_nets,
  names.arg = interventions$year,
  xlab = "Year",
  ylab = "Number of nets"
)
```

## Cases

The second example uses case data by year and age band.  For
simplicity we assume constant treatment coverage and diagnostic usage.

```{r case-data}
cases <- data.frame(
  year      = rep(1:6, each = 3),
  age_lower = rep(c(0, 5, 15), 6),
  age_upper = rep(c(5, 15, 100), 6),
  cases     = rep(rpois(6, 100), each = 3),
  tx_cov    = rep(c(0, 0.3, 0.8), each = 3),
  par       = rep(300, 6 * 3)
)
cases
```

We can estimate the number of rapid diagnostic tests (RDTs) and ACT
doses required using `commodity_rdt_tests()` and `commodity_al_doses()`.
The age specific dose multipliers are determined by the `age_upper`
values.

```{r commodities}
prop_rdt <- 1
prop_act <- 1

cases$n_rdt <- commodity_rdt_tests(
  n_cases            = cases$cases,
  treatment_coverage = cases$tx_cov,
  proportion_rdt     = prop_rdt
)

cases$n_act <- commodity_al_doses(
  n_cases            = cases$cases,
  treatment_coverage = cases$tx_cov,
  proportion_act     = prop_act,
  age_upper          = cases$age_upper
)

cases
```

Costs can then be added using the corresponding costing functions.

```{r case-costs}
cases$cost_rdt <- cost_rdt(cases$n_rdt)
cases$cost_act <- cost_al(cases$n_act)
head(cases)
```

```{r plot-cases, fig.cap="RDTs and ACT doses by year"}
barplot(
  rbind(cases$n_rdt, cases$n_act),
  beside = TRUE,
  names.arg = cases$year,
  legend.text = c("RDTs", "ACT doses"),
  xlab = "Year",
  ylab = "Number"
)
```

The examples above show the basic approach for moving from programme
data to estimates of commodity requirements and associated costs.  More
complex scenarios can be handled in the same way by adjusting the input
arguments to the commodity and costing functions.
